Demes: a standard format for demographic models

### 2022-08-01 response-to-reviewers.tex

\end{reply}

\begin{point}
I usually like when there is a link to the software directly in the abstract
 in those kind of papers, it avoids scrolling through it as links are
sometimes in different places in those type of manuscripts.
\end{point}
\begin{reply}

We agree that this is a useful practice.
However, in our case there are several software packages that are directly
and indirectly related to the Demes format as summarized in Table 1.
We have instead added a link to the specification documentation in the abstract.
This documentation includes a tutorial introduction for users to learn
how to write a Demes YAML file, and a detailed specification that developers
may consult when implementing Demes support in their own sofware.
\end{reply}

\begin{point}
I suggest including the example of the YAML format in the main text.
This is too sad that the center piece of this paper has been relegated to the appendix :(
Show us how easy it is to understand the model from the YAML format!
Maybe FigA2 is a bit too long, but FigA1 A and B could work well.
\end{point}
\begin{reply}
We thank the reviewer for the suggestion.
We have moved the example model in Figure A1.A (the HDM YAML) and
Figure A1.B (the model diagram) into the main text as Figures 1A and B,
while leaving the MDM form of the same model in the Appendix.
We have also added a section to the text that decribes the main features
of the model in relation to the YAML document.
\end{reply}

### 2022-08-01 paper.tex
Add section explaining Figure 1
Closes #77.

\ggcomment{Add some text here describing the YAML in Figure \ref{fig:IM}.}

@@

    \caption{
        \label{fig:IM}
        Example isolation-with-migration Demes model. (A) The Human Data Model
        representation expressed using YAML.
        (B) A visual representation of the model using \texttt{demesdraw}.
        The same model in the Machine Data Model form is provided in
        Figure~\ref{fig:IM-MDM}.
        The YAML descriptions in (A) and Figure~\ref{fig:IM-MDM} correspond
        exactly to a JSON description, but are much more human-readable.
    }
\end{figure}

>>>>

@@

    \caption{
        \label{fig:IM}
        Example isolation-with-migration Demes model. (A) The Human Data Model
        representation expressed using YAML.
        (B) A visual representation of the model using \demesdraw.
        The same model in the Machine Data Model form is provided in
        Figure~\ref{fig:IM-MDM}.
        The YAML descriptions in (A) and Figure~\ref{fig:IM-MDM} correspond
        exactly to a JSON description, but are much more human-readable.
    }
\end{figure}

In Figure \ref{fig:IM} we provide an example model in Demes YAML format.
Models typically start with a concise description, followed by the mandatory
\texttt{time\_units} field.
This model uses the \texttt{defaults} section to provide a default \texttt{start\_size}
of 1000 individuals for each epoch of each deme.
There are three demes in the model, an ancestral deme named ``A'' which ceases
to exist at 100 generations ago, and demes ``X'' and ``Y'' that derive their
ancestry from ``A'' at this time.
Demes A and X have only one epoch, in which the population sizes are
constant, whereas deme Y has two epochs. Deme Y's second epoch has a different
\texttt{end\_size} than its \texttt{start\_size}, which indicates the size
grows exponentially from 1000 individuals at 50 generations ago to 3000
individuals at time 0 (the present).
The migration section lists one migration stanza, between demes X and Y.
This migration stanza doesn't indicate a source or destination deme,
so the migration is symmetric. No migration times are specified, so migrations
occur continuously at the given rate during the time interval over which
both demes exist (from 100 generations ago until the present).
We do not attempt a detailed explanation of all Demes features here,
and readers are instead directed to the tutorial included
in the online Demes specification
(\url{https://popsim-consortium.github.io/demes-spec-docs/}).
